Prestige bias in cultural evolutionary dynamics

### 2023-03-01 ms/ms.tex
mwf suggestions

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{varioref}
\usepackage[unicode=true,colorlinks=true,urlcolor=blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
% appendix
\usepackage[title,page]{appendix}
\usepackage{chngcntr}
% unbreakable dashes
\usepackage[shortcuts]{extdash}
% footnotes
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

\begin{abstract}
A common bias in cultural transmission is success bias, where more successful individuals are more likely to be copied.
Because success may be difficult to estimate, individuals may indirectly estimate success by using prestige as a proxy for success. Here, we model prestige bias as a combination of indirect success bias and influence bias, the latter meaning that the choice of a role-model depends on the number of individuals that have already copied from that role-model.
We analyze the effect of prestige on cultural evolutionary dynamics using mathematical analysis and stochastic simulations.
We find analytic approximations to the stochastic role-model choice process, which facilitate further mathematical analysis and reduce the computational complexity of simulations. We also find approximations to the fixation probability and the fixation time of an invading cultural trait in different environments. 
We show that indirect success bias effectively plays the role of natural selection, whereas influence bias effectively plays the role of genetic drift.	
Influence bias also accelerates the evolutionary dynamics, as can be expected in a rich-getting-richer process.
Our study extends the literature that describes and analyses cultural transmission,
especially in human societies where social media is popular and thus influence bias may be expected.
\end{abstract}

@@

%\paragraph{Cultural transmission.}
In cultural transmission, individuals transmit cultural traits (i.e., behaviors, beliefs, norms) to one another, typically by learning and demonstrating \citep{transmissionVectorsBook}.
Examples for cultural traits in humans are norms, preferences, tools, ideas, and language, potentially transmitted verbally and/or by observations \citep{transmissionVectors}. 
Although cultural transmission is common in humans, it is also observed in other animals, e.g. chimpanzees \citep{chimpsPrestige, chimpsCopy}, dolphins and whales \citep{dolphins_whales}, elephants \citep{elepahntsRepo}, and even flies \citep{fliesPaper} and fish \citep{Truskanov2020}.

Cultural transmission may combine vertical transmission, in which parents transmit to their offspring; oblique transmission, in which adults (teachers, leaders, and even strangers) transmit to unrelated offspring; and horizontal transmission, in which peers from the same age cohort transmit to one another \citep{transmissionVectorsBook,transmissionVectors}.
It has been demonstrated that non-vertical cultural transmission can maintain maladaptive traits, which can be beneficial in changing environments \citep{changeEnv}.

Transmission biases may cause a cultural trait to have a disproportionate probability to be transmitted compared to its frequency in the population. \textit{Success bias} occurs when individuals prefer to copy from role models that demonstrate success in some activity, such as fishing, growing yams, using medicinal plants \citep{fijian_social_bias}, or hunting \citep{Mesoudi2008}, and it can increase the probability of learning a successful trait \citep{Borofsky2022}. 
Indeed, in a tournament between learning strategies conducted by \citep{strategiesPaper}, most winner strategies included a mixture of success-biased social learning and individual learning, implying that success-biased learning is a good strategy, but that by itself it is not enough to best other strategies, even when success is measured accurately.

% TODO edit so that we focus on indirect success - aka prestige - and decompose it to content based - which we call success - and context based - which we call influence
\citep[Ch.~5]{evolutionBook} have suggested that the evaluation of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
A direct bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can attempt the observed paddle grips to determine which grip is better.
A frequency-dependent bias occurs when the probability to copy a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{non-conformity} \citep{conformism}. 
The effects of conformity and non-conformity on cultural evolution have been studied with both models \citep{anticonformity,Denton2021,conformity} and experiments \citep{negativeFrequency}.
An indirect bias occurs when an individual uses one phenotype to determine the attractiveness of another phenotype.
For example, an observer may copy the paddle grip of the ping-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
This, however, may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, \citep[Ch. 8]{evolutionBook} have suggested that maladaptive traits may spread widely in a population if indirect biases are strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, \citep{cooperation} have shown that helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient, or when the benefit is much larger than the cost.

\citep{fijian_social_bias} have studied such indirect success biases, which they call cross-domain success bias or \emph{prestige bias} (e.g., great fishermen may be chosen as role-models for growing yams). They have suggested that such biases, over generations, can lead to cultural adaptations, and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits.
\citep[Ch.~8]{evolutionBook} and \citep{complexityPaper} have suggested that prestige biases are probably more common in humans than success biases, and prestige is often mentioned in the cultural evolution literature, however there are few models of prestige bias.

A broader definition of prestige is the ``widespread respect and admiration felt for someone or something on the basis of a perception of their achievements or quality'' (New Oxford American Dictionary).
Indeed, \citep{prestige_cultural_learning} have defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'', and demonstrated its occurrence in in 3-4 year old children.
\citep{prestige_evolution} gave a similar definition of prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature.

To distinguish this form of indirect bias from other definitions of prestige, we call it \emph{influence bias}.
This is an important distinction, as influence is a context-based bias, rather than a content-based bias: it does not depend on the phenotype itself but rather on the number of copiers that already copied each role-model, which may be easier and more accurate to estimate compared to success. 
Influence bias also differs from frequency biases such as conformity, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

In contemporary human society, social media makes it especially easy to estimate the social and cultural influence individuals have over others, and therefore can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices were invented to capitalize on this effect \citep{facebook_marketing}.

In the following, we develop a stochastic model of cultural transmission with prestige bias. In our model, prestige combines both cross-domain indirect success bias and influence bias.
We develop analytic approximations for this model and analyze its dynamics. We find approximations for the probability and time for fixation of a `successful' phenotype (i.e., that is favored by success bias). Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and influence bias play the role of natural selection and genetic drift, respectively.

@@

We begin with a continuous trait model with indirect bias suggested by \citep{evolutionBook}, develop an extension with influence bias, and then develop a model with a dichotomous trait.
We implemented our stochastic models and approximations, performed statistical analyses, and produced figures using Python \citep{python} with NumPy \citep{numpy} and Matplotlib \citep{mathplotlib}. 
Source code is available at \href{https://github.com/yoavram-lab/PrestigeBias}{https://github.com/yoavram-lab/PrestigeBias}.

\subsection*{Continuous trait}
We follow the model of \citep{evolutionBook}, assuming only oblique transmission of the trait and omitting the indirect trait in order to reduce model complexity. 
We consider a population of $N$ individuals, described by a single trait with a continuous value.
Each generation, $N$ naive individuals, or copiers, choose an individual from the previous generation, or role-models, from which they will copy their trait. Similar to a Wright-Fisher model, we assume non-overlapping generations such that the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t$. We assume the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
\citep[Ch.~8, p.~247-249]{evolutionBook} describe a blended transmission algorithm by defining $F$ as a weighted average of the traits of all role-models, 
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model,
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)} \;,
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$, equivalent to relative fitness in evolutionary-genetic transmission models.

\citep{evolutionBook} note that the deterministic blended transmission algorithm they use has alternatives. We thus develop a similar stochastic model with transmission from a single random role-model. Instead of \cref{eq:boydF}, we define the transmission function $F$ as a random variable with its distribution given by 
\begin{equation}\label{eq:ourF}
\Pr\big(F_i(\vec{A}) = A_{j}\big) = G_{i,j} \;,
\end{equation}
such that $G_{i,j}$ gives the probability of copier $i$ to choose to copy the trait of role-model $j$.

\paragraph{Influence bias.}
Here we introduce a new element to the model.
We assume that in each generation, copiers choose their role-models one by one.
We denote $K_{i,j}$ as the number of copiers that chose role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers already chose a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the prestige of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The prestige $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by success--the estimated biased trait value $\beta(A_{i,j})$--and influence--the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} of \citep{evolutionBook} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;.
\end{equation}

@@

We assume every individual may evaluate the importance of success and influence differently. Additionally, we assume every role-model displays his influence and success individually. For example, a person with more followers but lack of actual skill may emphasize the number of his followers rather than his actual skill. We therefore defined $\alpha_{i,j}$ to be the relation between every copier's intrinsic significance of estimation, and every role-model's personal emphasis.

@@

\subsection*{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, for simplicity and for easier mathematical analysis, we assume $\alpha$ is homogeneous, such that we have exchangeability between role-models.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + \alpha'X} \;,
\end{equation}
where $X$ is the number of role-models with trait $\hat{A}$, $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model, $\beta(\hat{A})=1$, and $\alpha'= \frac{\alpha}{1-\alpha}$ is the odd ratio of the bias weight. Complementing this, the probability of the $i$-th copier to copy phenotype $\hat{A}$ is $G_{i, \hat{A}} = 1-G_{i,A}$.
The rest of the details follow the continuous trait model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.
Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two such approximations, summarized here and explained in detail below. In both we assume either $\alpha_j$ is a trait of role-model $j$ only (contrary to $\alpha_{ij}$ that included the relation with copier $i$), or a completely homogeneous $\alpha$.

@@

\subsection*{Generalized binomial distribution}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

@@

\begin{proof}
We denote $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ as the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ our of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1) \;,
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citep{GBD}, which completes the proof.
\end{proof}
This result gives the following two corollaries on the expected number of followers of a given role-model by the end of the role-model choice process. The proof of \Cref{cor:GBD} is in Appendix~\ref{sec:GBD}.
\\

@@

\begin{proof}
\citep[eq.~(2.3)]{GBD} show that 
$\E[k] = N \cdot Q_j(1,1)$ (modified from their notation).
$Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made.
$Q_j(1,1) = G_{1,j}$ by definition, which completes the proof.
\end{proof}

@@

Although basic, \Cref{fig:GBD} shows good fit of the GB approximation.
This validation is initial, and the more extensive validations we do on the Dirichlet-Multinomial approximation, because it is what we will use in our analysts.

@@

>>>>

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{varioref}
\usepackage[unicode=true,colorlinks=true,urlcolor=blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
% appendix
\usepackage[title,page]{appendix}
\usepackage{chngcntr}
% unbreakable dashes
\usepackage[shortcuts]{extdash}
% footnotes
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

\begin{abstract}
{If the traits of more successful individuals are more likely to be adopted, the resulting cultural transmission is described as being success-biased. Prestige may be used as a proxy for successs.}
Here, we model prestige bias as a combination of indirect success bias and influence bias, the latter meaning that the choice of a role-model depends on the number of individuals that have already copied  that role-model.
{The} effect of prestige on cultural evolutionary dynamics {is analyzed} using mathematical analysis and stochastic simulations.
{Analytic} approximations to the stochastic role-model choice process {facilitate the} mathematical analysis and reduce the computational complexity of simulations. {Approximations are given}  to the fixation probability and the fixation time of an invading cultural trait in different environments. 
We show that indirect success bias effectively plays the role of natural selection, whereas influence bias effectively plays the role of genetic drift.	
Influence bias{, which may be strong in communities where social media are prevalent,}  also accelerates the evolutionary dynamics, as can be expected in a rich-{get}-richer process.
\end{abstract}

@@

%\paragraph{Cultural transmission.}
%In cultural transmission, individuals transmit cultural traits (i.e., behaviors, beliefs, norms) to one another, typically by learning and demonstrating \citep{transmissionVectorsBook}.
%Examples for cultural traits in humans are norms, preferences, tools, ideas, and language, potentially transmitted verbally and/or by observations \citep{transmissionVectors}. 
%Although cultural transmission is common in humans, it is also observed in other animals, e.g. chimpanzees \citep{chimpsPrestige, chimpsCopy}, dolphins and whales \citep{dolphins_whales}, elephants \citep{elepahntsRepo}, and even flies \citep{fliesPaper} and fish \citep{Truskanov2020}.

Cultural transmission {of attitudes, preferences, beliefs, norms, and behaviors} may combine vertical transmission, in which parents transmit to their offspring; oblique transmission, in which adults (teachers, leaders, and even strangers) transmit to unrelated offspring; and horizontal transmission, in which {individuals} from the same age cohort transmit to one another \citep{transmissionVectorsBook}.
It has been demonstrated that non-vertical cultural transmission can maintain maladaptive traits, which can be beneficial in changing environments \citep{changeEnv,Lehmann2009}.

Transmission biases may cause a cultural trait to have a {higher rate of transmission than} its frequency in the population. \textit{Success bias} occurs when individuals prefer to copy from role models that demonstrate success in some activity, such as fishing, growing yams, using medicinal plants \citep{fijian_social_bias}, or hunting \citep{Mesoudi2008}, and it can increase the probability of learning a {trait that is present in those successful individuals} \citep{Borofsky2022}. 
Indeed, in a tournament between learning strategies  \citep{strategiesPaper}, most {winning} strategies included a mixture of success-biased social learning and individual learning, implying that success-biased learning is a good strategy, but that by itself it is not enough to best other strategies, even when success is measured accurately.

% TODO edit so that we focus on indirect success - aka prestige - and decompose it to content based - which we call success - and context based - which we call influence
{Boyd and Richerson} \citep[Ch.~5]{evolutionBook}  suggested that the {assessment} of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
{Direct} bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can {try} the observed paddle grips to determine which grip is better.
{Frequency}-dependent bias occurs when the probability {of copying} a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{{anti}-conformity} \citep{conformism}. 
The effects of conformity and {anti}-conformity on cultural evolution have been studied with both models \citep{anticonformity,Denton2021,conformity} and experiments \citep{negativeFrequency}.
{Indirect} bias occurs when an individual uses one phenotype to determine the attractiveness of another phenotype.
For example, an observer may copy the paddle grip of the {ping}-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
{However, this} may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, {Boyd and Richerson} \citep[Ch. 8]{evolutionBook} {suggest} that maladaptive traits may spread widely in a population if indirect {bias is} strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient, or when the benefit is much larger than the cost \citep{cooperation}.

{Heinrich and Broesch \citep{fijian_social_bias}  studied} such indirect success biases, which they call cross-domain success bias or \emph{prestige bias} (e.g., great fishermen may be chosen as role-models for growing yams). They {suggested} that such biases, over generations, can lead to cultural adaptations, and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits
\citep[Ch.~8]{evolutionBook}. {Prestige bias may be} more common in humans than success {bias} \citep{complexityPaper}, and prestige is often mentioned in the cultural evolution literature, {although} there are few models of {it}.

A broader definition of prestige is the ``widespread respect and admiration felt for someone or something on the basis of a perception of their achievements or quality'' (New Oxford American Dictionary).
Indeed, {Chudek et al.}\ \citep{prestige_cultural_learning} have defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'', and demonstrated its occurrence in in 3-4 year old children.
{Henrich and Gil-White} \citep{prestige_evolution} gave a similar definition of prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature.

To distinguish this form of indirect bias from other definitions of prestige, we call it \emph{influence bias}.
This is an important distinction, as influence is a context-{dependent} bias, rather than a content-{dependent} bias: it does not depend on the phenotype itself but rather on the number of copiers that {have} already copied each role-model, which may be easier and more accurate to estimate {than} success. 
Influence bias also differs from {conformity}, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

In contemporary human society, social media {make} it especially easy to estimate the social and cultural influence individuals have over others, {which} can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices {have been} invented to capitalize on this effect \citep{facebook_marketing}.

In the following, we develop a stochastic model of cultural transmission with prestige bias {that} combines both cross-domain indirect success bias and influence bias.
We develop analytic approximations for this model and analyze its dynamics. We {also} find approximations for the probability and time {to} fixation of a `successful' phenotype (i.e., that is {subject to} success bias). Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and influence bias play the role of natural selection and genetic drift, respectively.

@@

We begin with a continuous trait model with indirect bias suggested by {Boyd and Richerson} \citep{evolutionBook}, {propose} an extension with influence bias, and then develop a model with a dichotomous trait.
We {implement} our stochastic models and approximations, {perform} statistical analyses, and {produce} figures using Python \citep{python} with NumPy \citep{numpy} and Matplotlib \citep{mathplotlib}. 
Source code is available at \href{https://github.com/yoavram-lab/PrestigeBias}{https://github.com/yoavram-lab/PrestigeBias}.

\subsection*{Continuous trait}
We follow the {Boyd and Richerson model} \citep{evolutionBook}, assuming only oblique transmission of the trait and omitting the indirect trait in order to reduce model complexity. 
{Consider} a population of $N$ individuals, described by a single trait {that takes continuous values}.
Each generation, $N$ naive individuals, or copiers, choose an individual from the previous generation, or role-models, from which they will copy their trait. Similar to a Wright-Fisher model, {generations are non-overlapping, and}  the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t${, and}  the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
{Boyd and Richerson} \citep[Ch.~8, p.~247-249]{evolutionBook} describe a blended transmission algorithm by defining $F$,  a weighted average of the traits of all role-models, {as}
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value {that} copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model,
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)} \;,
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$.

{Boyd and Richerson} \citep{evolutionBook} note that the deterministic blended transmission algorithm they use has alternatives. We {can} develop a similar stochastic model with transmission from a single random role-model{\ where instead} of \cref{eq:boydF}, we define the transmission function $F$ as a random variable with its distribution given by 
\begin{equation}\label{eq:ourF}
\Pr\big(F_i(\vec{A}) = A_{j}\big) = G_{i,j} \;;
\end{equation}
{here $G_{i,j}$ is} the probability {that copier $i$ chooses} to copy the trait of role-model $j$.

\paragraph{Influence bias.}
Here we introduce a new element to the model {by assuming} that in each generation, copiers choose their role-models one by one.
{Denote by} $K_{i,j}$  the number of copiers that {choose} role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers {had} already {chosen} a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the prestige of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The prestige $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by success--the estimated biased trait value $\beta(A_{i,j})$--and influence--the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;.
\end{equation}

@@

{Different individuals} may evaluate the importance of success and influence differently. Additionally, we assume {each} role-model displays {its} influence and success individually. For example, {individuals} with more followers but {lacking} skill may emphasize the number of {their} followers rather than {their} skill. %We therefore defined $\alpha_{i,j}$ to be the relation between every copier's intrinsic significance of estimation, and every role-model's personal emphasis.

@@

\subsection*{Dichotomous trait}
{We} introduce a simplified version where the trait {has only} two phenotypes: {an} optimal phenotype $\hat{A}$ and {a} sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability {that} that phenotype {is} transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, for simplicity and for easier mathematical analysis, we assume $\alpha$ is homogeneous, {which entails} exchangeability between role-models.
Therefore, the probability {that} the $i$-th copier {copies} phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + \alpha'X} \;,
\end{equation}
where $X$ is the number of role-models with trait $\hat{A}$, $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model, $\beta(\hat{A})=1$, and $\alpha'= \frac{\alpha}{1-\alpha}$ is the {odds} ratio of the bias weight. Complementing this, the probability of the $i$-th copier to copy phenotype $\hat{A}$ is $G_{i, \hat{A}} = 1-G_{i,A}$.
The rest of the details follow the continuous trait model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the {dynamics of the} phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.
Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two such approximations, summarized here and explained in detail below. In both we assume either $\alpha_j$ is a trait of role-model $j$ only ({different from} $\alpha_{ij}$ that included the relation with copier $i$), or a completely homogeneous $\alpha$.

@@

\subsection*{Generalized binomial distribution}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is {denoted}  by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

@@

\begin{proof}
{Let} $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ {be} the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ {out} of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1) \;,
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citep{GBD}, which completes the proof.
\end{proof}
This result gives the following two corollaries on the expected number of followers of a given role-model by the end of the role-model choice process. The proof of \Cref{cor:GBD} is in Appendix~\ref{sec:GBD}.
\\

@@

\begin{proof}
{Drezner and Farnum} \citep[eq.~(2.3)]{GBD} show that 
$\E[k] = N \cdot Q_j(1,1)$ (modified from their notation).
$Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made.
$Q_j(1,1) = G_{1,j}$ by definition, which completes the proof.
\end{proof}

@@

Although basic, \Cref{fig:GBD} shows good fit of the GB approximation.
This validation is initial, and {we do} more extensive validations on the Dirichlet-Multinomial approximation, because {that} is what we will use in our {analysis}.

@@

\bibitem{Lehmann2009}
Lehmann, L., and M.W. Feldman. Coevolution of adaptive technology, maladaptive culture, and population size in a producer-scrounger game. Proc.\ Roy.\ Soc.\ B 276 (2009): 3853-3862.

\bibitem{Fogarty2017}Fogarty, L., J. Y. Wakano, M. W. Feldman, and K. Aoki. The driving forces of cultural complexity: Neanderthals, modern humans, and the question of population size. Hum.\ Nat.\ (2017): 28: 39-52.
